AA: the Algorithmic Auto-regulation (Software Development) Methodology

### 2013-06-01 paper.tex
fixed tables size
revisado artigo todo na escrita

\indraftnote{
A very good article on the value of asynchronous communication for personal
and group productivity, related to the key necessity of having moments of
introversion to avoid daily pressures of forced socialization. The way we work
on the digital age enables people to be very productive, the article also
mentions Linux as a hallmark example~\cite{Thompson:Wired:2012}
}

\indraftnote{TODO: cite CIA.vc bot stuff}

\section{Related Work}
\label{related-work}

@@

There has been a large amount of research done in the area of methodologies to
deal with distributed teams of developers. We are focusing in GSD here, however
some principles involved on those methodologies could be used on smaller teams
of developers working in the same place, time and with minor cultural
differences. Moreover we generally think on ``distributed development'' being
global which is not totally true. We even applied AA to a team that work at the
same city but on different times (more details on Section~\ref{results}). Even
smaller groups of developers working on the same building could use GSD
methodologies. A thorough survey of these methodologies is beyond the scope of
this paper. Here we present a brief overview.

Various methodologies for GSD were built around the factors that affect
distributed team works, proposed by Carmel~\cite{carmel1999} and comprising
three distances: geographical, cultural and temporal. Geographical distance
prejudice \emph{coordination}, being the act of integrating all the tasks
distributed between units~\cite{carmel2001}; \emph{control}, or the process to
maintain specific goals, policies or quality levels; and
\emph{communication}. All these factors are correlated, for example, a team need
to have clearly communication to work on tasks of a specific problem. Cultural
distance encompass differences on organizational and natural culture. Spoken
language, unit and ethnic values are common forms of this distance. Some
companies prefer to situate development units in foreign locations with minimal
cultural distance (e.g.\ an American firm prefer Ireland, because of spoken
language similarity~\cite{carmel2001}). And finally the temporal distance that
prejudices synchronous communications like telephone or video-conferences. Units
of developers working on different time-zones are concerned with managing of
their agenda guided by this temporal distance.

Targeting geographical distance, Carmel~\cite{carmel2001} suggests a tactic to
reduce intensive collaboration. His approach divides the whole software
life-cycle on levels of complexity. Each level has a degree of
collaboration. For example, some developers working on a project with high
collaboration level should use the follow-the-sun approach: when concluding the
work day, they pass their work to the team working on another time-zone. Other
tactics are suggested by the same author to deal with the three distances, like
separating foreign units of developers in time-zone bands.

Battin et al.~\cite{battin2001} propose and argues about their experiments using
specific methodologies created for the distributed development centers of the
Motorola Company (which has over 25 software development centers
worldwide). These methodologies included constant communication with critical
units, incremental integration and scheduled based on time-zones distributed to
developers on 6 countries from 3 continents.

While considering free software projects instead of companies, the same factors
are present and some methodologies arises. German~\cite{german2003} gives a
concise review of methodologies used by the GNOME project, one of the most
active free software projects and used by companies like Sun Microsystems. It is
interesting to note a difference on viewpoint present on his paper: German
focuses the methodology description on code. He start explaining that GNOME is
separated into modules (76 on version 2.4, to be precise) and each module has
one maintainer who divide his modules into separated parts in which other
developers can work on independent tasks, along other responsibilities. As like
others free software projects, all the development was done using a
\emph{bugtracker} to bugs and issues management, mail lists and Internet Relay
Chat (IRC) to discussion and communication and a software configuration
management like SVN or Git. Periodical (commonly yearly) conferences like GUADEC
is common on free and open source projects to face-to-face meeting and is based
each time on a different place.

\section{The AA Methodology}
\label{aa-methodology}

As noted, some strategies for GSD is based on complex methodologies and many of
those were built for a specific company or software center. Here we propose an
alternative methodology based on a simple idea: small working sessions logged by
a computational tool. Figure~\ref{fig:mm} summarizes our methodology.

\begin{figure}
\begin{center}
   \includegraphics[width=0.8\linewidth,keepaspectratio=true]{figs/aa-mm.png}
\end{center}
   \caption{
   Mindmap of our methodology
   }
\label{fig:mm}
\end{figure}

\subsection{AA Session}

From the developer perspective, the AA methodology is based on creating pretty
small high perspective reports of what they are doing in a specific time frame,
that can be something between 5 to 15 minutes, depending in what is more
comfortable for the developer. An AA Session would be a period of at least 2
hours doing these reports. The developer can set reminders to show up when its
time to make a new report. The objective of the flexible time frame and
reminders is to minimize developer overhead during his AA session. In this way
he can make the reports while staying concentrated in his code. Each report can
be sent directly to an on-line server, or stored locally in a temporarily
database for sending later. This make possible the use of AA while offline.

Developers can also record a video screencast in the end of the session
summarizing what has been done, explaining with his words and showing his most
important results. This, combined with the textual log of his session, makes the
whole report more understandable for himself or other people searching for
information about his production.

\subsection{AA Website Report}

\begin{figure}
\begin{center}
   \includegraphics[width=0.95\linewidth]{figs/aa-0_1.png}
\end{center}
   \caption{AA Version 0.1}
\label{fig:aaserver}
\end{figure}

All AA reports made by the developers are sent to a Web server and become public
in a Website, being possible then for a manager or another developer to follow
very closely the work of a developer, nearly real-time, reading each of his
small reports of what he is doing right now.

Another possibility is to check older sessions to see when sometimes was done
and the comments of the developer about it, since each AA post happens in a very
small time frame of work, the information about what was done become very easy
to understand, instead of a long report in the end of a session.

The site is where the developer can add to his session a link for his screencast
about that session, a video summary of what was done to complement the reports,
useful specially on cases where the small reports were done in a hurry, because
the developer did not want to lose his focus on something important at that
moment.

\subsection{Validation}

Each AA session must be validated by another developer, it means that all
reports are read by someone that will consider then valid or not and will even
write commentaries about the specific session. The developer in charge of
validating a session is decided randomly by the AA Web server, which send an
email to the chosen developer with an URL to a validation interface.

% TODO: adicionar figura aa cliente (ou tabela com os principais comandos)
% TODO: adicionar figura com arquitetura do AA cliente + AA web + validação
% TODO: melhorar screenshot do AA web

\section{Results and Discussion}
\label{results}

The easy and effective management of teams working on GSD is the main purpose of
the AA methodology. We applied this methodology to a group of 9 developers in
July of 2011 on what we called Lab Macambira~\footnote{LabMacambira.sf.net:
  \url{http://labmacambira.sf.net}}. The main objective of the team was to work
on different free software projects, contributing directly to their development,
sending bug corrections or proposing new features on their source code.

All the team members had different levels of knowledge on software development
as part of large and distributed free software projects like Scilab or
Mozilla. In this way, one month of training was conducted by two experienced
developers, teaching the basics of use of development support tools like
bugtrackers, programming languages and version control systems. After this
period, a challenge was proposed for the new developers: send a bug correction
or a new feature to a large free software project and you will work with us, you
will be a ``Macambira'' developer. Table~\ref{tabela:contribuicoes} summarizes
the contributions of each ``Macambira'' to free software projects.

@@

\begin{table}
  \caption{Free and open software projects that received contributions
    from ``Macambiras''. On the first column we can see a list of
    applications of those projects. At right, the pseudonym of the
    ``Macambiras'' who sent \textit{commits} to the application. At
    ``Lab Macambira'', and at free software community in general, is a
    common practice to use pseudonym as identification.}
    \small\begin{tabular}{|l|l|}
        \hline
        Application           & ``Commiters''                       \\
        \hline \hline
        Mozilla Firefox       & daneoshiga, bzum                    \\
        Evince                & hick209, bzum, marcicano, mquasar   \\
        BePDF / Xpdf          & marcicano                           \\
        Ekiga                 & flecha                              \\
        Empathy               & fefo                                \\
        Lib Folks (Telepathy) & kamiarc                             \\
        Scilab                & v1z, humannoise                     \\
        VxL                   & v1z                                 \\
        ImageMagick           & v1z                                 \\
        OpenOffice            & hick209                             \\
        Puredata              & v1z, automata, greenkobold, gilson, bzum \\
        Puredata OpenCV       & v1z                                 \\
        Puredata GEM          & v1z, fefo, hick209                  \\
        Puredata PDP          & v1z, fefo, hick209                  \\
        ChucK                 & rfabbri, automata                   \\
        ChucK MiniAudicle     & rfabbri, automata                   \\
        WebRTC                & automata                            \\
        OSC-Web               & automata                            \\
        Web-PD-GUI            & automata                            \\
        Live-Processing       & automata                            \\
        ChucK-Wiimote         & automata                            \\
        Audiolet              & automata                            \\
        Extempore             & automata                            \\
        \hline
    \end{tabular}
    \label{tabela:contribuicoes}
\end{table}

In one month, each developer contributed to many very large free software
projects. Many of the developers started the training with no knowledge of what
is free software and ended that period becoming a free software developer.

During that month, the same team developed the first version of the AA system
and used AA to manage their activities. Even while developing the system. All
the source code of AA --- the client that sends the logs and the AA Web server
--- is public available~\footnote{AA source code:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
and all the AA sessions log of the whole team of ``Lab Macambira'' is also
on-line~\footnote{Logs of AA sessions:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.

After the training period, during more 6 months, the ``Macambiras'' worked on a
large range of free software projects, distributed on work groups --- each work
group focusing on a specific theme like video, audio and web --- and financed by
contracts and support of the ``Pont\~{a}o N\'{o}s Digitais''. In
Table~\ref{tabela:criados} we can see a list of the free software created by
``Lab Macambira'' since July of 2011.

\begin{table}
    \caption{Software projects created by ``Lab Macambira'' since July
      of 2011 with a short description and the technologies --- like
      programming languages or frameworks --- involved. It is
      interesting to note the heterogeneity of projects and its areas
      of application.}
    \begin{tabular}{|l|p{5cm}|l|}
        \hline
        Application & Description & Technologies involved \\ 
        \hline \hline
        AA            & Algorithmic Auto-regulation      & Python, PHP \\
        \hline
        \'{A}gora Communs & System for on-line deliberations & PHP \\
        \hline
        SIP           & Scilab Image Processing toolbox & C, Scilab \\
        \hline
        animal        & An Imaging Library              & C \\
        \hline
        TeDi          & Test Framework for Distance Transform
        Algorithms & C, Shellscript, Scilab \\
        \hline
        Macambot      & Multi-use IRC Bot               & Python \\
        \hline
        ``Confer\^{e}ncia Permanente'' & Platform for the permanent
        conference of the rights of minors & PHP, JavaScript \\
        \hline
        CPC           & Center for accounting of the Brazilian culture
        representation groups & Python, Django \\
        \hline
        Timeline      & Interactive time lines on the Web & JavaScript
        \\
        \hline
        Imagemap      & Interactive marking for on-line photos &
        JavaScript \\
        \hline
        ABT           & Program for real-time execution and musical
        rhythmic analysis & Python \\
        \hline
        EKP           & Emotional Kernel Panic & Python, ChucK \\
        \hline
        SOS           & Aggregation and diffusion of popular and native
        knowledge about health & Python, Django \\
        \hline
        Creative Economy & Platform for creative, collaborative and
        solidarity economy of the culture hubs and cultural entities &
        Python, Django \\
        \hline
        OpenID Integration & Adaptations to existing software for
        unified login through OpenID & PHP \\
        \hline
        pAAinel & Dashboard for the real-time visualization of Lab
        Macambira activity & Python, Django \\
        \hline
        Georef & Collection of scripts to be used as reference, which
        aims to be a GIS platform to map public data of use to
        citizens & Python, Django \\
        \hline
        AirHackTable & Software for an instrument which generates
        sound from flying origami tracked by webcams & Puredata,
        C/C++, Scilab \\
        \hline
        \end{tabular}
    \label{tabela:criados}
\end{table}

@@

%% brief introduction
In a scenario where GSD is growing as a popular form of software development,
not just on free software projects but in the whole software industry, we need
methodologies to deal with its disadvantages and at the same time to amplify its
advantages.

This paper has presented a methodology to GSD, being the development conducted
on large or small groups of software developers, working on different countries
or even at the same room. The AA methodology implements a simple system where
each developer take notes of his work generating a periodical log of small text
sentences. The sum of those sentences, along an entire session of work, results
in a complete report. The report is made public available through a Website and
be validated by other developers sorted aleatory by the AA Web server.

Instead of a merely work-management tool, AA act as a methodology to improve the
time sense of individuals, dividing their work on small sessions, and also
reducing the need of extensive reports or unnecessary meetings. By asking users
to write a minimal text sentence as a continuously log, AA does not disturb
developers concentrated on programming: developers just have to type some
characters, hit \textit{enter} and go back to coding.

AA application is not restricted to software development. As of this writing
there is a comic book studio~\footnote{Pula pirata: \url{http://pulapirata.com}}
starting to use AA to manage their activities. People with non-software
background, like social scientists, musicians and activists has also using AA
and contributing for its improvement.

For developer teams, we have experienced the use of AA to auto-regulate the work
of ``Lab Macambira'', a group of free software developers from Brazil. Since
July of 2011 the group have contributed and created new free and open source
software for a vast number of applications.

There are many aspects of the work which remain unfinished. New ways to report
logs --- the ``Twitter like'' messages --- from different interfaces like IRC,
Internet Messaging services and email can make the use of AA easy and
widespread, turning AA an ubiquitous system, presented on everyday communication
channels. Even the work logs generated since July of 2011 could be vastly
statistically analyzed aiming to recognize patterns in the behavior of
individuals and their productions.

We would like to conclude setting an important role of AA: being a free software
system and an open methodology, AA could be used to auto-manage groups of
individuals working on software or other kinds of activities. In this way, we
are interested to spread AA for those groups, to have even more developers
contributing in a collaborative way.

@@

We would like to also thank AA: the present research and even this manuscript
was written using AA. The complete log is on-line at
\url{http://www.pulapirata.com/skills/aa}.

>>>>

\section{Related work}
\label{related-work}

@@

There has been a large amount of research in methodologies to
deal with distributed teams of developers. Although this article focuses in GSD,
some principles involved on those methodologies could be used on smaller teams
of developers working in the same place, time period and with minor cultural
differences. Moreover, ``distributed development'' is generally tough as being
global which is not true. AA was applyed to a team that work at the
same city but on different times (more details on Section~\ref{results}). Even
smaller groups of developers working on the same building could use GSD
methodologies. A thorough survey of these methodologies is beyond the scope of
this paper. This section presents a brief overview.

Various methodologies for GSD were built around the factors that affect
distributed team works. As proposed by Carmel~\cite{carmel1999}, these comprises
three distances: geographical, cultural and temporal. Geographical distance
prejudice \emph{coordination}, being the act of integrating all the tasks
distributed between units~\cite{carmel2001}; \emph{control}, or the process to
maintain specific goals, policies or quality levels; and
\emph{communication}. All these factors are correlated, for example, a team needs
to have good communication to work on tasks related to a specific problem. Cultural
distance encompass differences on organizational and natural culture. Spoken
language, unit and ethnic values are common forms of this distance. Some
companies prefer to situate development units in foreign locations with minimal
cultural distance (e.g.\ an American firm prefer Ireland, because of spoken
language similarity~\cite{carmel2001}). Finally, the temporal distance
prejudices synchronous communications like telephone or video-conferences.
Developers working on different time-zones manage
their agenda guided by this temporal distance.

Targeting geographical distance, Carmel~\cite{carmel2001} suggests a tactic to
enhance intensive collaboration. His approach divides the whole software
life-cycle on levels of complexity. Each level has a degree of
collaboration. For example, some developers working on a project with high
collaboration level should use the follow-the-sun approach: when concluding the
work day, they pass their work to the team working on another time-zone. Other
tactics are suggested by the same author to deal with the three distances, like
separating foreign units of developers in time-zone bands.

Battin et al.~\cite{battin2001} argues about their experiments using
methodologies created for the development centers of the
Motorola Company (which has over 25 software development centers
worldwide). These methodologies included constant communication with critical
units, incremental integration and schedules based on time-zones distributed to
developers on 6 countries from 3 continents.

While considering free software projects instead of companies, similar factors
are present and some methodologies arises. German~\cite{german2003} gives a
concise review of methodologies used by the GNOME project, one of the most
active free software projects and used by companies like Sun Microsystems. It is
interesting to note a difference on the viewpoint present on his paper: German
focuses the methodology description on code. He start explaining that GNOME is
separated into modules (76 on version 2.4, to be precise) and each module has
one maintainer who divide his modules into separated parts in which other
developers can work on independent tasks, along other responsibilities. As like
others free software projects, all the development was done using a
\emph{bugtracker} to bugs and issues management, mail lists and Internet Relay
Chat (IRC) to discussion and communication and a software version
management like SVN or Git. He finishes by observing that periodical, e.g. yearly, conferences like GUADEC
is common on free and open source projects for face-to-face meetings and is based
each time on a different place.

\section{The AA methodology}
\label{aa-methodology}

As noted, some strategies for GSD is based on complex methodologies and many of
those were built for a specific company or software center. This section proposes an
alternative based on a simple idea: small working sessions logged by
a computational tool. Figure~\ref{fig:mm} summarizes this methodology.

\begin{figure}
\begin{center}
   \includegraphics[width=0.8\linewidth,keepaspectratio=true]{figs/aa-mm2.png}
\end{center}
   \caption{
   Mindmap of AA methodology: 1) developer engagement cycle, i.e. his use of AA; 2) functionality as main goals of the system; 3) potentialities as enhancements AA delivers to the developer's medium. 
   }
\label{fig:mm}
\end{figure}

\subsection{The AA session}

From a developer's perspective, the AA methodology is based on creating
small, high perspective, reports of what is being done in a specific time frame. This frame
can be something between 5 to 15 minutes, depending on what is 
comfortable for the developer. An AA Session is a period of about 2
hours, in which these short reports (called ``shouts'') are periodically written. The developer can set reminders to show up when it is
time to make a new report. The objective of the flexible time frame and
reminders is to minimize developer overhead during his AA session. In this way,
he can make the reports with little or no loss of concentration on his coding activity. Each report can
be sent directly to an online server, or stored locally in a temporary
database for sending later on. This makes it possible to use AA while offline.

Developers can also record a video screencast at the end of the session,
summarizing what has been done, explaining with his words and showing on the screen his most
important results. This, combined with the textual log of his session, makes the
whole report more significant as documentation and understandable for himself or other people searching for
information about his production.

\subsection{The AA website report}

\begin{figure}
\begin{center}
   \includegraphics[width=0.95\linewidth]{figs/aa-0_1_.png}
\end{center}
   \caption{AA Version 0.1. Messages of users: hybrid, filter0, v1z e aut0mata about activities of interest for labMacambira.sf.net and other parties (IPRJ/UERJ, IFT/UNESP, IFSC/USP, OPW/Mozilla, PulaPirataComics). Each message is a ``shout'', which, grouped, forms an AA session.}
\label{fig:aaserver}
\end{figure}

All AA reports made by the developers are sent to a Web server and become public
in a Website. It is then possible for a manager or another developer to follow
closely the work of a developer, nearly real-time, reading each of his
small reports of what he is doing.

Another possibility is to check older sessions to see what was done
and the comments of the developer about it. Since each AA post (or ``shout'') happens in a
small time frame of work, the information about what was done becomes easy
to understand, which seems better than a long report in the end of a session.

The developer can add to his session a link for his screencast
about that session. This screencast is a video summary of what was done and complements the report.

\subsection{Validation}
Each AA session must be validated by another developer. This means that all
reports are read by someone that will consider valid or not and will even
write comments about the specific session. The developer in charge of
validating a session is decided randomly by the AA Web server, which sends an
email to the chosen developer with the URL to a validation interface.

% TODO: adicionar figura aa cliente (ou tabela com os principais comandos)
% TODO: adicionar figura com arquitetura do AA cliente + AA web + validação
% TODO: adicionar figura do pAAinel

\section{Results and discussion}
\label{results}

The easy and effective management of teams working on GSD is the main purpose of
the AA methodology. This methodology was applied to a group of 9 developers in the second half of 2011 on what was named ``Lab Macambira''\footnote{LabMacambira.sf.net:
  \url{http://labmacambira.sf.net}}. The main objective of the team was to work
on different free software projects, contributing directly to their development,
sending bug corrections or proposing new features on their source code.

Team members earned knowledge on software development
as part of large and distributed free software projects like Scilab and  
Mozilla. In this way, one month of training was conducted by three experienced
developers, teaching the basics of development support tools like
bugtrackers, programming languages and version control systems. After this
period, a challenge was proposed for the new developers: send a bug correction
or a new feature to a large free software project and you will work with us, you
will be a ``Macambira'' developer. Table~\ref{tabela:contribuicoes} summarizes
the contributions of each ``Macambira'' to free software projects.

@@

\begin{table}
  \caption{Free and open software projects that received contributions
    from ``Macambiras''. On the first column there is a list of
    applications. At right, the pseudonym of the
    ``Macambiras'' who sent \textit{commits} to the application. At
    ``Lab Macambira'', and at free software community in general, is a
    common practice to use a pseudonym as identification.}
    \begin{center}
    \small\begin{tabular}{|l|l|}
        \hline
        Application           & ``Commiters''                       \\
        \hline \hline
        Mozilla Firefox       & daneoshiga, bzum                    \\
        Evince                & hick209, bzum, marcicano, mquasar   \\
        BePDF / Xpdf          & marcicano                           \\
        Ekiga                 & flecha                              \\
        Empathy               & fefo                                \\
        Lib Folks (Telepathy) & kamiarc                             \\
        Scilab                & v1z, humannoise                     \\
        VxL                   & v1z                                 \\
        ImageMagick           & v1z                                 \\
        OpenOffice            & hick209                             \\
        Puredata              & v1z, automata, greenkobold, gilson, bzum \\
        Puredata OpenCV       & v1z                                 \\
        Puredata GEM          & v1z, fefo, hick209                  \\
        Puredata PDP          & v1z, fefo, hick209                  \\
        ChucK                 & rfabbri, automata                   \\
        ChucK MiniAudicle     & rfabbri, automata                   \\
        WebRTC                & automata                            \\
        OSC-Web               & automata                            \\
        Web-PD-GUI            & automata                            \\
        Live-Processing       & automata                            \\
        ChucK-Wiimote         & automata                            \\
        Audiolet              & automata                            \\
        Extempore             & automata                            \\
        \hline
\end{tabular}
\end{center}
    \label{tabela:contribuicoes}
\end{table}

In one month, ``Macambira'' developers contributed to many large free software
projects. Some of the developers started the training with no knowledge of what
is free software and ended that period becoming a free software developer.

During that month, the same team developed the first version of the AA system
and used AA to manage their activities, even while developing the system. All
the source code of AA --- the client that sends the logs and the AA Web server
--- is publicly available\footnote{AA source code:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
and all the AA session logs of the ``Lab Macambira'' team is also
online\footnote{Logs of AA sessions:
  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.

After the training period, during 6 more months, the ``Macambiras'' worked on a
large range of free software projects, distributed on work groups. Each work
group focused on a specific theme like video, audio and web. They were financed by
contracts and support of the ``Pont\~{a}o de Cultura Digital: N\'{o}s Digitais''. In
Table~\ref{tabela:criados} we can see a list of the free software created by
``Lab Macambira'' since July of 2011.

\begin{table}
    \caption{Software projects, created by ``Lab Macambira'', with the use of AA, since July
      of 2011 with a short description and the technologies --- like
      programming languages or frameworks --- involved. It is
      interesting to note the heterogeneity of projects and its areas
      of application.}
    \small\begin{tabular}{|l|p{5cm}|l|}
        \hline
        Application & Description & Technologies involved \\ 
        \hline \hline
        AA            & Algorithmic Auto-regulation      & Python, PHP \\
        \hline
        \'{A}gora Communs & System for on-line deliberations & PHP \\
        \hline
        SIP           & Scilab Image Processing toolbox & C, Scilab \\
        \hline
        animal        & An Imaging Library              & C \\
        \hline
        TeDi          & Test Framework for Distance Transform
        Algorithms & C, Shellscript, Scilab \\
        \hline
        Macambot      & Multi-use IRC Bot               & Python \\
        \hline
        ``Confer\^{e}ncia Permanente'' & Platform for the permanent
        conference of the rights of minors & PHP, JavaScript \\
        \hline
        CPC           & Center for accounting of the Brazilian culture
        representation groups & Python, Django \\
        \hline
        Timeline      & Interactive time lines on the Web & JavaScript
        \\
        \hline
        Imagemap      & Interactive marking for on-line photos &
        JavaScript \\
        \hline
        ABT           & Program for real-time execution and musical
        rhythmic analysis & Python \\
        \hline
        EKP           & Emotional Kernel Panic & Python, ChucK \\
        \hline
        SOS           & Aggregation and diffusion of popular and native
        knowledge about health & Python, Django \\
        \hline
        Creative Economy & Platform for creative, collaborative and
        solidarity economy of the culture hubs and cultural entities &
        Python, Django \\
        \hline
        OpenID Integration & Adaptations to existing software for
        unified login through OpenID & PHP \\
        \hline
        pAAinel & Dashboard for the real-time visualization of Lab
        Macambira activity & Python, Django \\
        \hline
        Georef & Collection of scripts to be used as reference, which
        aims to be a GIS platform to map public data of use to
        citizens & Python, Django \\
        \hline
        AirHackTable & Software for an instrument which generates
        sound from flying origami tracked by webcams & Puredata,
        C/C++, Scilab \\
        \hline
        \end{tabular}
    \label{tabela:criados}
\end{table}

@@

%% brief introduction
In a scenario where GSD is growing as a popular form of software development,
not just on free software projects but in the whole software industry, there is an increased need for
methodologies to deal with it's disadvantages and at the same time to amplify it's
advantages.

This paper has presented a simple methodology to GSD, being the development conducted
on large or small groups of software developers, working on different countries
or even at the same room. The AA methodology implements a simple system where
each developer takes notes of his work, generating a periodical log of small text
sentences. The sum of those sentences, along an entire session of work, results
in a complete report. The report is made public available through a Website and
are validated by other developers sorted aleatory by the AA Web server.

Instead of a merely work-management tool, AA acts as a methodology to improve the
time sense of individuals, dividing their work on small sessions, and also
reducing the need of extensive reports or unnecessary meetings. By asking users
to write a minimal text sentence as a continuous log, AA minimizes
disturbances: developers just type some
characters, hit \textit{enter} and go back to coding, which can help mental organization about the work being done.

AA application is not restricted to software development. As of this writing
there is a comic book studio\footnote{Pula Pirata Comics: \url{http://pulapirata.com}}
using AA to manage their activities. People with non-software
background, like social scientists, musicians and activists has also been using AA
and contributing for its improvement.

For software developer teams, the use of AA was used successfully to auto-regulate the work
of ``Lab Macambira'', a group of free software developers from Brazil. Since
July of 2011 the group have contributed and created new free and open source
software for a vast number of applications.

There are many aspects of the work which remain unfinished. New ways to report
logs --- the ``Twitter like'' messages, sometimes called ``shouts'' --- from different interfaces like IRC,
Internet Messaging services and email can make the use of AA easy and
widespread, turning AA an ubiquitous system, presented on everyday communication
channels. Even the work logs generated since July of 2011 could be 
statistically analyzed to recognize patterns in the behavior of
individuals and their productions.

Being a free software
system and an open methodology, AA could be used to auto-manage groups
working on software or other kinds of activities. In this way,
there is interest on spreading AA to have even more developers
contributing in a collaborative way.

@@

Authors thank AA: the present research and even this manuscript was written using AA. The complete log is on-line at
\url{http://www.pulapirata.com/skills/aa}.
Authors are also grateful to IFSC/USP, IPRJ/UERJ, IFT/UNESP and all AA users and collaborators, spacially those who coded AA hacks for shell and bot uses and those who coded the different Web interfaces in current use.
